### 2020-12-13 paper/main.tex
kill a HOGG; tweaks

% to-do
% -----
% - Make sure that what we say about kriging and Wiener are correct.
% - Are we clear that the feature weighting is a kind of regularization? I (DWH) think it IS because it affects how the L2 on beta affects different parameters. SV says that maybe we should get explicit about the regularization points, and when the regularization is implicit or hidden.
% - In uncertainties, we talk about Bayesian. Should we also talk about frequentist and/or information theory?
% - Audit for the use of "points" and "locations" and "data" to make sure all is consistent. The ts are locations, the ys are data.
% - Audit for use of the words "fit" and "prediction" and "model". All okay?
% - Audit for use of \, as the multiply operator? Fix the kerning around the ^\top items?
% - Audit for design matrix, feature matrix, and so on. We should mention design matrix but stick with feature matrix as our terminology.
% - Audit for hyphens in under-parameterization and re-weighting and so on? Hyphenated words or single words? Make decisions and stick to them.

@@

% Price-Whelan's figure hacks
\usepackage[framemethod=tikz]{mdframed}
\usetikzlibrary{shadows}
\definecolor{captiongray}{HTML}{555555}
\mdfsetup{%
  innertopmargin=2ex,
  innerbottommargin=1.8ex,
  linecolor=captiongray,
  linewidth=0.5pt,
  roundcorner=5pt,
  shadow=false,
  shadowcolor=black!05,
  shadowsize=4pt
}

@@

\section*{\raggedright Fitting very flexible models: Linear regression with large numbers of parameters%
\footnote{%
It is a pleasure to thank
Jed Brown (CU Boulder),
Dan Foreman-Mackey (Flatiron),
Teresa Huang (JHU),
Sam Roweis (deceased),
Adrian Price-Whelan (Flatiron),
Bernhard Sch\"olkopf (MPI-IS), and
Rachel Ward (UT Austin)
for valuable conversations and input.
SV is partially funded by NSF DMS 2044349, EOARD FA9550-18-
1-7007, and the NSF-Simons Research Collaboration on the Mathematical and Scientific
Foundations of Deep Learning (MoDL) (NSF DMS 2031985)}}

@@

HOGG: Now do the full generalization with $C$ and $\Lambda$ and explain how the matrix inversion formula (CITE SOMETHING) makes these two expressions equivalent:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top\,C^{-1}\,X + \Lambda)^{-1}\,X^\top\,C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top\,(X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
Either leave the proof to the reader or make it an exercise HOGG: Check the units. And continue discussion of connections to Bayesian inference. Explain that these forms are good because they relieve the fits of the burden of going exactly through every data point. This is more responsive to the real situation of noisy data, and permits smoother solutions.
HOGG: AND return to units again. And cite products again.

@@

Given a data set (location--data pairs $t_i, y_i$), what is the best way to empirically estimate, from those data, the out-of-sample prediction error---that is, how do you estimate how well you are likely to predict new data?
The only good answer to this is cross-validation:
In cross-validation, we leave out a part of the data (in the most extreme form, leave out one single data point at a time), train the model using all but the left-out part, and predict the left-out part. Then this process is iterated over all choices of what part (or point) to leave out.
This process is illustrated in \figurename~\ref{fig:loo}, where we show $23$ fits, each of which is trained to the 22 points left when we leave one of the $n=23$ points out.
Also shown is the prediction, in each leave-one-out fit, for the left-out point.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/LOO.pdf}
    \caption{A demonstration of leave-one-out cross-validation for one particular model. The 23 lines are the 23 fits, for each of which one data point was held out. Also ploted on each of the 23 lines is the prediction made for that fit's held-out data point. HOGG anything else to say here?}
    \label{fig:loo}
    \end{mdframed}
\end{figure}

The prediction for the mean-squared error (MSE) for new data is the mean of the square of the differences between the leave-one-out predictions and the left-out data.
This leave-one-out cross-validation MSE (CVMSE) will be different for different choices for the basis, size ($p$), and regularization of the fits you do.
It is a fairly reliable and well-studied method for assessing predictive accuracy (HOGG: WHAT TO CITE HERE?), although it does depend on some assumptions (for example, about the distribution of predictive information across individual data points, maybe).
An example of CVMSE for two models (the feature-weighted OLS and the ridge regression) are shown for our toy data, as a function of the number of features $p$, in \figurename~\ref{fig:cv}.
The predictive accuracy is a strong function of $p$!
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/cross-validation.pdf}
    \caption{Leave-one-out cross-validation estimates of mean squared prediction error for the feature-weighted OLS fits, as a function of the number of features $p$. Also shown are the same for ridge regression with $\lambda=0.1$ and also the Gaussian process introduced in \sectionname~\ref{sec:gp} and shown in \figurename~\ref{fig:gp}. The GP result is shown as a flat line because it doesn't have an associated number $p$. In detail the first (lowest-$t$) and last (highest-$t$) data points were not used in computing the mean squared error; that choice is debatable, but we are imagining an assessment of the quality of the \emph{interpolations}, not extrapolations, of these models. Note that the OLS predictions are much, much worse at $p\approx n$ (check out all the orders of magnitude on the vertical axis) than they are at very low or very high $p$, but that ridge regression doesn't show this behavior.}
    \label{fig:cv}
    \end{mdframed}
\end{figure}

@@

Importantly for us, there is a strong connection between the feature-weighted OLS and the GP.
In particular, when the features $x_j$ form some kind of (SOLEDAD: WHAT TO SAY HERE?) basis, and we take the limit of infinite features ($p\rightarrow\infty$) we get kernel matrices in place of the $X\,\Lambda^{-1}\,X^\top$ matrix products;
\begin{align}
    \lim_{p\rightarrow\infty} X\,\Lambda^{-1}\,X^\top &= K
    \\
    \lim_{p\rightarrow\infty} X_\ast\,\Lambda^{-1}\,X^\top &= K_\ast
    ~,
    \label{eq.limit}
\end{align}
where $\Lambda^{-1}$ is the diagonal matrix of weights, and element $i,i'$ of $K$ is obtained by evaluating a kernel function $k(t_i,t_{i'})$.
Equivalently, the limit is
\begin{equation}
    \lim_{p\rightarrow\infty} \sum_{j=1}^p [\Lambda^{-1}]_{jj}\,g_j(t)\,g_{j}(t') = k(t, t')
    ~,
\end{equation}
where we have used the diagonality of $\Lambda$ to make a single sum over $j$.
The specific form of the kernel function $k(\cdot,\cdot)$ depends on the basis (the features) we choose, and the weighting of the basis functions in the OLS.

@@

At the end of \sectionname~\ref{sec:fwols}, we discussed including not just feature weights in a matrix $\Lambda^{-1}$ but also data weights in a matrix $C^{-1}$. The GP also permits this, and it is often a very good idea. The generalization of equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2} to the GP case is 
\begin{equation}\label{eq:gpC}
    \hat{Y}_\ast = K_\ast\,(K + C)^{-1}\,Y
    ~.
\end{equation}
As before, in contexts where you have independent uncertainties on each element $y_i$ of your training data $Y$, $C^{-1}$ would naturally be set to the diagonal matrix containing the inverses of the variances of those uncertainties (so $C$ would contain the variances).
This form \eqref{eq:gpC} is used a lot in astronomy and physics (HOGG CITE SOME THINGS?) and it is also the standard form given for the Wiener filter, where the kernel function generating $K$ is the Fourier tranform of the ``signal power'' and the kernel function generating $C$ is the Fourier transform of the ``noise power''.

@@

You might think that all of this is academic, but it really isn't when the number of features $p$ is within a factor of a few of the number of data points $n$.
For example, in our toy-data OLS experiments in this paper with $n=23$, the matrix $X^\top\,X$ has a condition number (ratio of largest eigenvalue to smallest nonzero eigenvalue) that saturates machine precision for the entire range $12 < p < 50$!

@@

We only considered the setting in which you care about predicting new data $Y_\ast$ and never the setting in which you care about the internal parameters $\beta$ of the linear fit.
However, if you are, say, measuring the power spectrum of a process (as we do, often, in cosmology), then you care about these amplitudes themselves.
In this case, any regularization you apply takes the role of a prior or prior information.
That prior must be chosen (in those cases) with great care, because the answers you get will be influenced by that choice; when it doesn't properly conform to your true beliefs, your answers will be distorted in bad ways.
And, technically, when you take this view (that the parameters $\beta$ matter), you also have to get serious about the noise model on the data.
That is, the beliefs encoded in the data noise variance tensor $C$ also must represent your true beliefs if you don't want your answers distorted in bad ways.
All of that is out of scope here, but we say things about it all elsewhere (\citealt{fitting}).

@@

In our toy examples, we used a Fourier series.
That is just one choice among many.
However, \emph{it is often a great choice}.
When you choose the Fourier basis (and---importantly---for every cosine term you include the corresponding sine term), the matrix $X\,\Lambda^{-1}\,X^\top$ (and the limiting kernel matrix $K$ at $p\rightarrow\infty$) has the property that every matrix element $[X\,\Lambda^{-1}\,X^\top]_{ii'}$ depends (or can be calculated from) just the absolute difference $|t_i-t_{i'}|$.
That is, in this basis, all fitting methods become technically \emph{stationary}.
This all relates to the translation-independence properties of the Fourier basis.
So although the Fourier basis is just one choice among many, it is the \emph{right choice} when you think your problem has (or might have) certain kinds of translation invariances.

@@

\appendix
\section{Mathematical arguments and results}

>>>>

% to-do
% -----
% - Make sure that what we say about Wiener filtering are correct.
% - Are we clear that the feature weighting is a kind of regularization? I (DWH) think it IS because it affects how the L2 on beta affects different parameters. SV says that maybe we should get explicit about the regularization points, and when the regularization is implicit or hidden.
% - In uncertainties, we talk about Bayesian. Should we also talk about frequentist and/or information theory?
% - Audit for the use of "points" and "locations" and "data" to make sure all is consistent. The ts are locations, the ys are data.
% - Audit for use of the words "fit" and "prediction" and "model". All okay?
% - Audit for use of \, as the multiply operator? Fix the kerning around the ^\top items?
% - Audit for design matrix, feature matrix, and so on. We should mention design matrix but stick with feature matrix as our terminology.
% - Audit for hyphens in under-parameterization and re-weighting and so on? Hyphenated words or single words? Make decisions and stick to them.

@@

% Price-Whelan's figure hacks
\usepackage[framemethod=tikz]{mdframed}
\usetikzlibrary{shadows}
\definecolor{captiongray}{HTML}{555555}
\mdfsetup{%
  innertopmargin=2ex,
  innerbottommargin=1.8ex,
  linecolor=captiongray,
  linewidth=0.5pt,
  roundcorner=1pt,
  shadow=false,
  %shadowcolor=black!05,
  %hadowsize=4pt
}

@@

\section*{\raggedright Fitting very flexible models: Linear regression with large numbers of parameters%
\footnote{%
It is a pleasure to thank
Jed Brown (CU Boulder),
Dan Foreman-Mackey (Flatiron),
Teresa Huang (JHU),
Sam Roweis (deceased),
Adrian Price-Whelan (Flatiron),
Bernhard Sch\"olkopf (MPI-IS), and
Rachel Ward (UT Austin)
for valuable conversations and input.
SV is partially funded by NSF DMS 2044349, EOARD FA9550-18-
1-7007, and the NSF-Simons Research Collaboration on the Mathematical and Scientific
Foundations of Deep Learning (MoDL) (NSF DMS-2031985).}}

@@

In the most general case, you have not just a set of feature weights $\Lambda^{-1}$, you also have a set of data-point weights $C^{-1}$ (which, in standard settings, would be the inverses of the variances of the noise affecting the data points $y_i$).
When you put these all together, the feature-weighted, data-weighted least squares precitions are given by either of these two equivalent expressions:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top\,C^{-1}\,X + \Lambda)^{-1}\,X^\top\,C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top\,(X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
We discuss the equivalence of these below in \appendixname~\ref{app:math}. These expressions appear in Bayesian-inference contexts (see, for example \citealt{products}) when $\Lambda$ is the variance of the prior on the parameter vector $\beta$ and $C$ is the variance of the noise on the data vector $Y$. In those contexts, the units of $C$ would be the units of $Y$ squared, and the units of $\Lambda$ would be the square of the ratio of the units of $Y$ to the units of $X$ (the square of the units of $\beta$. These weighted-feature, weighted-data least-square forms are good because (if the weightings are chosen appropriately) they lead to smooth solutions that are not required to pass precisely through every data point. This is appropriate in common, real situations in which data are noisy and the world is smooth. Which form you choose, between \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}, depends on a few things, but primarily $p/n$. If $p<n$ it's both faster and more stable to use \eqref{eq:LambdaC1}; if $p>n$ it's faster and more stable to use \eqref{eq:LambdaC2}.

@@

Given a data set (location--data pairs $t_i, y_i$), what is the best way to empirically estimate, from those data, the out-of-sample prediction error---that is, how do you estimate how well you are likely to predict new data?
The only good answer to this is cross-validation:
In cross-validation, we leave out a part of the data (in the most extreme form, leave out one single data point at a time), train the model using all but the left-out part, and predict the left-out part. Then this process is iterated over all choices of what part (or point) to leave out.
This process is illustrated in \figurename~\ref{fig:loo}, where we show $23$ fits, each of which is trained to the 22 points left when we leave one of the $n=23$ points out.
Also shown is the prediction, in each leave-one-out fit, for the left-out point.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/LOO.pdf}
    \caption{A demonstration of leave-one-out cross-validation for one particular model. The 23 lines are the 23 fits, for each of which one data point was held out. Also ploted on each of the 23 lines is the prediction made for that fit's held-out data point. Some of the predictions don't appear within the plot window, because the predictions obtain large amplitudes at the edges.}
    \label{fig:loo}
    \end{mdframed}
\end{figure}

The prediction for the mean-squared error (MSE) for new data is the mean of the square of the differences between the leave-one-out predictions and the left-out data.
This leave-one-out cross-validation MSE (CVMSE) will be different for different choices for the basis, size ($p$), and regularization of the fits you do.
It is a fairly reliable and well-studied method for assessing predictive accuracy (HOGG: WHAT TO CITE HERE?), although it does depend on some assumptions (for example, about the distribution of predictive information across individual data points, maybe).
An example of CVMSE for two models (the feature-weighted OLS and the ridge regression) are shown for our toy data, as a function of the number of features $p$, in \figurename~\ref{fig:cv}.
The predictive accuracy is a strong function of $p$!
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/cross-validation.pdf}
    \caption{Leave-one-out cross-validation estimates of mean squared prediction error for the feature-weighted OLS fits, as a function of the number of features $p$. Also shown are the same for ridge regression with $\lambda=0.1$ and also the Gaussian process introduced in \sectionname~\ref{sec:gp} and shown in \figurename~\ref{fig:gp}. The GP result is shown as a flat line because it doesn't have an associated number $p$. In detail the first (lowest-$t$) and last (highest-$t$) data points were not used in computing the mean squared error; that choice is debatable, but we are imagining an assessment of the quality of the \emph{interpolations}, not extrapolations, of these models. Note that the OLS predictions are much, much worse at $p\approx n$ (check out all the orders of magnitude on the vertical axis) than they are at very low or very high $p$, but that ridge regression doesn't show this behavior. (HOGG MAKE SURE THIS POINT APPEARS IN THE MAIN TEXT).}
    \label{fig:cv}
    \end{mdframed}
\end{figure}

@@

Importantly for us, there is a strong connection between the feature-weighted OLS and the GP.
In particular, when the features $x_j$ form some kind of (SOLEDAD: WHAT TO SAY HERE? reproducing? Hilbert? ...?) basis, and we take the limit of infinite features ($p\rightarrow\infty$) we get kernel matrices $K$ in place of the $X\,\Lambda^{-1}\,X^\top$ matrix products:
\begin{align}
    \lim_{p\rightarrow\infty} X\,\Lambda^{-1}\,X^\top &= K
    \\
    \lim_{p\rightarrow\infty} X_\ast\,\Lambda^{-1}\,X^\top &= K_\ast
    ~,
    \label{eq.limit}
\end{align}
where $\Lambda^{-1}$ is the diagonal matrix of weights, and element $i,i'$ of $K$ is obtained by evaluating a kernel function $k(t_i,t_{i'})$.
Equivalently, the limit is
\begin{equation}
    \lim_{p\rightarrow\infty} \sum_{j=1}^p [\Lambda^{-1}]_{jj}\,g_j(t)\,g_{j}(t') = k(t, t')
    ~,
\end{equation}
where we have used the diagonality of $\Lambda$ to make a single sum over $j$.
The specific form of the kernel function $k(\cdot,\cdot)$ depends on the basis (the features) we choose, and the weighting of the basis functions in the OLS.

@@

At the end of \sectionname~\ref{sec:fwols}, we discussed including not just feature weights in a matrix $\Lambda^{-1}$ but also data weights in a matrix $C^{-1}$. The GP also permits this, and it is often a very good idea. The generalization of equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2} to the GP case is 
\begin{equation}\label{eq:gpC}
    \hat{Y}_\ast = K_\ast\,(K + C)^{-1}\,Y
    ~.
\end{equation}
As before, in contexts where you have independent uncertainties on each element $y_i$ of your training data $Y$, $C^{-1}$ would naturally be set to the diagonal matrix containing the inverses of the variances of those uncertainties (so $C$ would contain the variances).
This form \eqref{eq:gpC} is used a lot in astronomy and physics (HOGG CITE SOME THINGS?) and it is also the standard form given for the Wiener filter, where the kernel function generating $K$ is the Fourier transform of the ``signal power'' and the kernel function generating $C$ is the Fourier transform of the ``noise power''.

@@

You might think that all of this is academic, but it really isn't when the number of features $p$ is within a factor of a few of the number of data points $n$.
For example, in our toy-data OLS experiments in this paper with $n=23$, the matrix $X^\top\,X$ has a condition number (ratio of largest eigenvalue to smallest nonzero eigenvalue) that saturates machine precision for the entire range $12 < p < 50$.

@@

We only considered the setting in which you care about predicting new data $Y_\ast$ and never the setting in which you care about the internal parameters $\beta$ of the linear fit.
However, if you are, say, measuring the power spectrum of a process (as we are, often, in cosmology), then you care about these amplitudes themselves.
In this case, any regularization you apply takes the role of a prior or prior information.
That prior must be chosen (in those cases) with great care, because the answers you get will be influenced by that choice; when it doesn't properly conform to your true beliefs, your answers will be distorted in bad ways.
And, technically, when you take this view (that the parameters $\beta$ matter), you also have to get serious about the noise model on the data.
That is, the beliefs encoded in the data noise variance tensor $C$ also must represent your true beliefs if you don't want your answers distorted in bad ways.
All of that is out of scope here, but we say things about it all elsewhere (\citealt{fitting}).

@@

In our toy examples, we used a Fourier series.
That is just one choice among many.
However, \emph{it is often a great choice}.
When you choose the Fourier basis (and---importantly---for every cosine term you include the corresponding sine term), the matrix $X\,\Lambda^{-1}\,X^\top$ (and the limiting kernel matrix $K$ at $p\rightarrow\infty$) has the property that every matrix element $[X\,\Lambda^{-1}\,X^\top]_{ii'}$ depends on (or can be calculated from) just the absolute difference $|t_i-t_{i'}|$.
That is, in this basis, all fitting methods become technically \emph{stationary}.
This all relates to the translation-independence properties of the Fourier basis.
So although the Fourier basis is just one choice among many, it is the \emph{right choice} when you think your problem has (or might have) certain kinds of translation invariances.

@@

\appendix
\section{Mathematical arguments and results}\label{app:math}
